---
title: "Examples"
author: "Albert Y. Kim"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    df_print: default
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Set R chunk default options:
# -do not output all R messages in html report
# -do not output all R warnings in html report
# -set default figure widths and heights for all code blocks, unless you 
#  override them
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2
)
```

```{r, echo=FALSE}
# Load all packages here. It's good practice to load all packages at the
# beginning of a Rmd file so that collaborators can see in one shot all packages
# they'll need installed:
library(tidyverse)
library(fivethirtyeight)
library(leaflet)
library(sf)
library(maps)

# USA specific data
library(USAboundaries)
library(tidycensus)
library(tmaptools)
```



# Tip {-}

Do not run entire code blocks at once, as you'll skip over a lot of very important details. Rather, I recommend you highlight small segments of code and run them in the console individually using:

* macOS: Command + Enter
* Windows: Control + Enter



# 1. Converting data frames to sf objects

```{r}
# Convert pres_2016_trail data frame to sf object using st_as_sf() from sf
# package. Note we set the coordinate reference system to WGS84
# https://spatialreference.org/ref/epsg/wgs-84/
pres_2016_trail_sf <- pres_2016_trail %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326)
pres_2016_trail_sf

# Load maps package map object of US states and plot using default R plot
# function (not ggplot2). Plot is not pretty!
USA_map <-  maps::map("state", plot = FALSE, fill = TRUE)
plot(USA_map)

# Convert maps package map object to sf object using st_as_sf() again:
USA_sf <- USA_map %>% 
  st_as_sf(crs = 4326)

# Plot both layers:
ggplot() +
  geom_sf(data = USA_sf) +
  geom_sf(data = pres_2016_trail_sf, aes(col = candidate), size = 2) +
  facet_wrap(~candidate)
```


## Exercise

Draw a map of the US with the 4 colleges I've worked at marked:

```{r}
bert_colleges <- tibble(
  name = c("Reed College", "Middlebury College", "Amherst College", "Smith College"), 
  lat = c(45.481, 44.004, 42.370, 42.316), 
  lon = c(-122.631, -73.173, -72.516, -72.640)
)


```

Draw a map of only Massachusetts with some notable cities marked. This will require you to manually create a data frame like `bert_colleges` above. 

```{r}


```



# 2. Loading shapefiles into R

Let's load a shapefile of all bike trails in Massachusetts as provided by
[MassGIS (Bureau of Geographic Information)](https://docs.digital.mass.gov/dataset/massgis-data-bicycle-trails){target="_blank"}

```{r}
# Load the "shapefile" contained in the biketrails_arc/ folder and load
# it into R using the sf package's read_sf() function.
#
# Notice how the filepath is just "biketrails_arc/" and not 
# "~/Documents/SDS192/MP3/biketrails_arc/"
#
# In other words, R can find and load this shapefile independent of where the
# MP3 project folder is on your computer, as long are you are in MP3 RStudio
# Project mode
bike_trails <- "biketrails_arc/" %>%
  read_sf()

# Plot this data!
ggplot() +
  geom_sf(data = bike_trails, col = "red")

# The above map lacks context however. Where are these bikepaths?
# Let's obtain a map of Massachusetts counties from the USAboundaries package
MA_counties <- 
  USAboundaries::us_counties(resolution = "low", states = c("massachusetts")) %>% 
  select(-state_name)
  
# Plot this data!
ggplot() +
  geom_sf(data = MA_counties)

# Plot both maps by adding both layers!
ggplot() +
  geom_sf(data = bike_trails, col = "red") + 
  geom_sf(data = MA_counties) 

# Where did our bike paths go? The order you plot in matters! Switch the two
# geom_sf()
ggplot() +
  geom_sf(data = MA_counties) +
  geom_sf(data = bike_trails, col = "red")

# Better! But what are these counties? Let's mark each counties' centroid
MA_counties <- MA_counties %>%
  # Add centroids to each region using purrr package (covered in SDS 270 I believe)
  mutate(
    lon = purrr::map_dbl(geometry, ~st_centroid(.x)[[1]]),
    lat = purrr::map_dbl(geometry, ~st_centroid(.x)[[2]])
    )
MA_counties

# Plot! Notice how we define data = in the geom_point() instead of the ggplot().
# This sets the data set only for the geom_point() and not the other geoms.
# Any data = defined in the global ggplot() trickles down by default to the other
# geoms.
ggplot() +
  geom_sf(data = MA_counties, size = 0.5) +
  geom_sf(data = bike_trails, col = "red") +
  geom_point(data = MA_counties, aes(x=lon, y=lat)) 

# Can we do better? Yes! Use geom_text() instead with the label aesthetic!
ggplot() +
  geom_sf(data = MA_counties, size = 0.5) +
  geom_sf(data = bike_trails, col = "red") +
  geom_text(data = MA_counties, aes(x = lon, y = lat, label = name)) 
```


## Exercise

```{r}

```



# 3. Choropleth maps using census data

Note that the following code chunk has `eval = FALSE` for now so that R Markdown
does not evaluate it (i.e. skips it) for now. We'll flip this to `eval = TRUE`
later.

```{r, eval=FALSE}
# Get MA county sf object, that includes population estimate column in data
mass_pop_orig <- 
  tidycensus::get_acs(
    geography = "county", 
    variables = "B01003_001", 
    state = "MA",
    geometry = TRUE
  ) %>% 
  # Add centroids to each region using purrr package (I have no idea how this 
  # works!)
  mutate(
    lon = purrr::map_dbl(geometry, ~st_centroid(.x)[[1]]),
    lat = purrr::map_dbl(geometry, ~st_centroid(.x)[[2]])
  )  
mass_pop_orig

# Unfortunately the original NAME variable above is very long and redundant.
# Create a new data frame with the just the county name using the separate()
# function from the tidyr package. 
mass_pop <- mass_pop_orig %>% 
  tidyr::separate(NAME, c("County", "State"), sep = ", ") %>% 
  tidyr::separate(County, c("County", "Fluff"), sep = " ") 

# Plot gradient choropleth map based on numerical variable estimate
ggplot() +
  geom_sf(data = mass_pop, aes(fill = estimate), size = 0.5) +
  geom_text(data = mass_pop, aes(x=lon, y=lat, label = County)) +
  labs(fill = "Population\nEstimate")

# What if instead we want a bin based choropleth map. Use the cut_number()
# function to convert the numerical variable estimate to a categorical 
# variable with 3 levels!
mass_pop <- mass_pop %>%
  mutate(estimate_discrete = cut_number(estimate, 3))
mass_pop

# Plot!
ggplot() +
  geom_sf(data = mass_pop, aes(fill = estimate_discrete), size = 0.5) +
  geom_text(data = mass_pop, aes(x=lon, y=lat, label = County)) +
  labs(fill = "Population\nEstimate")
```


## Exercise

```{r}

```



# 4. Interactive maps using leaflet

```{r}
# Interactive map example:
leaflet() %>%
  addTiles() %>% 
  addMarkers(lng=-72.6419889, lat=42.3177692)

# How did I get the latitude and longitude for Smith College? By tapping into
# the OpenStreetMap (OSM) database using the geocode_OSM() function from the
# tmaptools package. Be careful though, apparently you are limited to 2000
# queries a day.
locations <- tibble(
  college = c("Reed College", "Middlebury College", "Amherst College", "Smith College")
)
locations_lat_long <- tmaptools::geocode_OSM(locations$college) %>% 
  as_tibble()

# lat and lon are the coordinates. I'm fairly sure all _min and _max values are
# the bounding box of the map shown when you search for a particular query in
# OpenStreetMap
locations_lat_long
```


## Exercise

```{r}

```


