---
title: "Problem Sets"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message=FALSE, warning = FALSE, eval = TRUE, 
  fig.width = 16/2, fig.height = 9/2
)

# Set seed value of random number generator to get "replicable" random numbers.
# Why 76? Because of https://www.youtube.com/watch?v=xjJ7FheCkCU
set.seed(76)
```

<!--
{target="_blank"}
-->

<style>
h1{font-weight: 400;}
</style>



***



# Extension Requests

* You have an extension "budget" of a maximum of 5 days for the rest of the semester; it's up to you to keep track of your budget
* Request must be made before due date/time.
* [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSeCrhqs-iTD5qPcf4FfcFNHE_rwMuaeF37yQuwusRsP9xHJ_Q/viewform){target="_blank"}



***



# Problem set 07 {#PS07}

*Due Monday 12/6 9pm*

For this problem set you will start building a GitHub portfolio of work. You need to complete all steps to receive full credit.



## 1. Edit your public GitHub profile

**Learning Goal**: Edit your public GitHub profile (if you haven't already). Think of this as a professional profile like a LinkedIn profile. Refer to Sec02 in-class data assistant Marium Tapal's [profile](https://github.com/mariumtapal){target="_blank"} as an example.

1. Add your name
1. Change your default profile picture. It doesn't have to include your face if you're not comfortable.
1. Add whatever other info to your profile you feel comfortable being public.


## 2. Create your first repo with a README

**Learning Goal**: Build your first README. Think of this as a cover page to market your project.

1. Get started
    1. Download the following zip file: <a href="static/PS/PS07.zip" download>`PS07.zip`</a>.
    1. Move `PS07.zip` to your SDS192 folder on your computer
    1. Unzip `PS07.zip`. Windows users: be sure to "Extract all"
    1. In the resulting PS07 folder, double-click the RStudio Project `PS07.Rproj` icon. For this PS you should **always** work in "PS07" RStudio Project mode
1. Change the default contents of `README.Rmd` to anything you like. In particular look at RStudio menu bar -> Help -> Markdown Quick Reference for markdown formatting tips.
    1. Title
    1. Section names
    1. The plot (change it to be a `ggplot2` plot)
1. Knit `README.Rmd` to produce `README.md` which is your repo's cover page
1. Put your work on GitHub. See screencast:

<iframe width="560" height="315" src="https://www.youtube.com/embed/4-9z_MIGvXY" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


## 3. Submission

Submit a link to this GitHub repository using this [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSf8wZIviqbDpf0jtecTGgg35LnKTvnW26tDZMhh1OB59bFexQ/viewform){target="_blank"}. 





<!--
1. **Start this problem set early!** That way you will have time to get help at the Spinelli Center tutoring hours if you run into trouble.
1. **I will not answer your questions until you've been to the Spinelli Center Sun-Thurs 7-9pm tutoring hours first.**
    + It will be very difficult for me to help 68 students individually. Please help me spread the workload. If you're stuck, please get help at the Spinelli Center first: that way they can handle the easy questions, leaving only the harder ones for me. I will be asking you if you've already gone to the Spinelli Center first before answering any of your questions.
    + Note that tutoring from 6pm on Tuesday November 26th through noon on Sunday December 1st is cancelled.
1. **GitHub frustrations are normal.** Git and GitHub are extremely powerful tools, but they can be very frustrating at times. This is normal. Check out the title of this [zine](https://gumroad.com/l/oh-shit-git){target="_blank"} made by [Julia Evans](https://twitter.com/b0rk?ref_src=twsrc%5Egoogle%7Ctwcamp%5Eserp%7Ctwgr%5Eauthor){target="_blank"} for advanced users.

## 1. GitHub component {#PS07_R}

**Note**: I sent an email on Saturday inviting you to the [Fall 2019 SDS192 Intro to Data Science
](https://github.com/192-2019-09-Fall){target="_blank"} GitHub Classroom Organization. Please accept this invitation.

For this PS you will be using the ["Happy Git and GitHub for the useR"](https://happygitwithr.com/){target="_blank"} book to:

1. Get familiar with GitHub. Read:
    + Chapter 1: Why Git? Why GitHub?
1. Install GitHub on your computers and set it up. Read and perform all steps in:
    + ~~Chapter 4: Register a GitHub account~~ You've already done this
    + ~~Chapter 5: Install or upgrade R and RStudio~~ You've already done this
    + Chatper 6: Install Git
    + Chapter 7: Introduce yourself to Git
1. Locate the GitHub repository you created in Lec32. 
    + You should be able to see it [here](https://github.com/192-2019-09-Fall){target="_blank"} and it should be named something like `learning-github-YOUR_GITHUB_USER_NAME`. So for example, my repo would be called `learning-github-rudeboybert`.
    + If you don't see it, accept this [invitation](https://classroom.github.com/a/sgKoAnsq){target="_blank"} to receive one. 
1. "Clone" the above repo into RStudio on your computer. Read and perform all steps in:
    + Chapter 12: Connect RStudio to Git and GitHub. **For Section 12.2** ignore what's in the book and do the following:
        1. You will not "Make a repo on GitHub," but rather use the `learning-github-YOUR_GITHUB_USER_NAME` repo from the previous step. 
        1. Click the big green button "Create repository."
        1. Copy the HTTPS clone URL to your clipboard via the green “Clone or Download” button.
1. You do not need to submit anything. Completing Chapter 12 will automatically update your submission on GitHub classroom.  


## 2. Hint {#PS07hint}

How do I know if I've successfully completed the assignment?

1. Go to the [SDS192 Intro to Data Science](https://github.com/192-2019-09-Fall) GitHub organization and make sure you see your `learning-github-YOUR_GITHUB_USER_NAME` folder.
1. Then go to your `learning-github-YOUR_GITHUB_USER_NAME` folder and then look at your README. There should be the line "This is a line from RStudio" on it. 
-->



***



# Problem set 06 {#PS06}

*Due Monday 11/15 9pm*

## 1. R Markdown component {#PS06_R}

Getting started:

* Download the following zip file: <a href="static/PS/PS06.zip" download>`PS06.zip`</a>.
* Move `PS06.zip` to your SDS192 folder on your computer
* Unzip `PS06.zip`. Windows users: be sure to "Extract all"
* In the resulting PS06 folder, double-click the RStudio Project `PS06.Rproj` icon. For this PS you sound **always** work in "PS06" RStudio Project mode (See Lec26 Mon 11/8)

Submission:

* You will only be submitting the resulting `PS06.pdf` file on gradescope. Please use the "Knit to PDF" method directly, instead of "Knit to Word" and then save to PDF method.
* If you encounter issues knitting your document, consult the [6 R Markdown Fixes](https://docs.google.com/document/d/1P7IyZ4On9OlrCOhygFxjC7XhQqyw8OludwChz-uFd_o/){target="_blank"}.
* Submit on [gradescope](https://www.gradescope.com/courses/308310){target="_blank"}:
    1. For each questions, tag **all** pages of your PDF that relate to that question (not just the first)
    1. You will be asked to tag the honor code question as well.



***



# Problem set 05 {#PS05}

*Due Monday 10/25 9pm*

## 1. R Markdown component {#PS05_R}

* Download the following Rmd template file: <a href="static/PS/PS05.Rmd" download>`PS05.Rmd`</a>.
* You will be submitting the resulting `PS05.pdf` file on gradescope. Please use the "Knit to PDF" method directly, instead of "Knit to Word" and then save to PDF method.
* Complete the individual questions. Knit early, knit often.
* If you encounter issues knitting your document, consult the [6 R Markdown Fixes](https://docs.google.com/document/d/1P7IyZ4On9OlrCOhygFxjC7XhQqyw8OludwChz-uFd_o/){target="_blank"}.
* Submit on [gradescope](https://www.gradescope.com/courses/308310){target="_blank"}:
    1. For each questions, tag **all** pages of your PDF that relate to that question (not just the first)
    1. You will be asked to tag the honor code question as well.



## 2. Solutions {#PS05_solutions}

* <a href="static/PS/PS05_solutions.Rmd" download>`PS05_solutions.Rmd`</a> R Markdown "source code" 
* [`PS05_solutions.pdf`](static/PS/PS05_solutions.pdf){target="_blank"} PDF


**Solutions to PS05 Q1.c)**. What happened to the average age above 60? The hint given on Slack was to look at the `ggplot2` cheatsheet -> 2nd page -> Bottom right corner -> "Zooming"

* `ylim(a, b)` sets the limit on the y-axis to be between `a` to `b` and "clips" (throws out) any points outside this interval
* `coord_cartesian(ylim=c(a, b))` zooms in on the the y-axis to be between `a` to `b` but does not "clip" (throw out) the points outside this interval

For example, consider the following regression line: 

```{r, fig.width = 8/1.5, height = 4.5/1.5, eval=TRUE}
library(ggplot2)
library(dplyr)

example <- tibble(
  x = c(1, 2, 3, 4, 5),
  y = c(1, 1, 1, 1, 5)
)

ggplot(example, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Let's set the y-axis limit to be between 0 and 3. Using `ylim(0, 3)` clips out the point `(5, 5)` and thus the regression line is flat:

```{r, fig.width = 8/1.5, height = 4.5/1.5, eval=TRUE}
ggplot(example, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ylim(0, 3)
```

However, using `coord_cartesian(ylim = c(0, 3))` merely zooms in on this part of the y-axis without clipping the point `(5, 5)` and thus the regression line is the original one: 

```{r, fig.width = 8/1.5, height = 4.5/1.5, eval=TRUE}
ggplot(example, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(0, 3))
```




***




# Problem set 04 {#PS04}

*Due Monday 10/18 9pm*

## 1. R Markdown component {#PS04_R}

* Download the following Rmd template file: <a href="static/PS/PS04.Rmd" download>`PS04.Rmd`</a>.
* You will be submitting the resulting `PS04.pdf` file on gradescope. Please use the "Knit to PDF" method directly, instead of "Knit to Word" and then save to PDF method.
* Complete the individual questions. Knit early, knit often.
* If you encounter issues knitting your document, consult the [6 R Markdown Fixes](https://docs.google.com/document/d/1P7IyZ4On9OlrCOhygFxjC7XhQqyw8OludwChz-uFd_o/){target="_blank"}.
* Submit on [gradescope](https://www.gradescope.com/courses/308310){target="_blank"}:
    1. Anytime between Wed 10/13 8pm and Mon 10/18 9pm
    1. For each questions, tag **all** pages of your PDF that relate to that question (not just the first)
    1. You will be asked to tag the honor code question as well.



## 2. Solutions {#PS04_solutions}

* <a href="static/PS/PS04_solutions.Rmd" download>`PS04_solutions.Rmd`</a> R Markdown "source code" 
* [`PS04_solutions.pdf`](static/PS/PS04_solutions.pdf){target="_blank"} PDF





***




# Problem set 03 {#PS03}

*Due Tuesday 9/28 9pm*

## 1. R Markdown component {#PS03_R}

* Download the following Rmd template file: <a href="static/PS/PS03.Rmd" download>`PS03.Rmd`</a>.
* You will be submitting the resulting `PS03.pdf` file on gradescope. Please use the "Knit to PDF" method directly, instead of "Knit to Word" and then save to PDF method.
* Complete the individual questions. Knit early, knit often.
* If you encounter issues knitting your document, consult the [6 R Markdown Fixes](https://docs.google.com/document/d/1P7IyZ4On9OlrCOhygFxjC7XhQqyw8OludwChz-uFd_o/){target="_blank"}.
* Submit on [gradescope](https://www.gradescope.com/courses/308310){target="_blank"}:
    1. Anytime between Mon 9/20 5pm and Tue 9/29 9pm
    1. For each questions, tag **all** pages of your PDF that relate to that question (not just the first)
    1. You will be asked to tag the honor code question as well.



## 2. Solutions {#PS03_solutions}

* <a href="static/PS/PS03_solutions.Rmd" download>`PS03_solutions.Rmd`</a> R Markdown "source code" 
* [`PS03_solutions.pdf`](static/PS/PS03_solutions.pdf){target="_blank"} PDF




***




# Problem set 02 {#PS02}

*Due Monday 9/20 5pm*

## 1. R Markdown component {#PS02_R}

* Download the following Rmd template file: <a href="static/PS/PS02.Rmd" download>`PS02.Rmd`</a>.
* You will be submitting the resulting `PS02.pdf` file on gradescope. There are two methods to create a PDF:
    1. "Knit to PDF". This is the preferred format (for consistency for the graders). 
    1. If this doesn't work for you, then "Knit to Word" and then export/save as PDF
* Complete the individual questions. Knit early, knit often.
* If you encounter issues knitting your document, consult the [6 R Markdown Fixes](https://docs.google.com/document/d/1P7IyZ4On9OlrCOhygFxjC7XhQqyw8OludwChz-uFd_o/){target="_blank"}.
* Submit on [gradescope](https://www.gradescope.com/courses/308310){target="_blank"}:
    1. Anytime between Mon 9/13 5pm and Mon 9/20 5pm
    1. For each questions, tag **all** pages of your PDF that relate to that question (not just the first)



## 2. Solutions {#PS02_solutions}

* <a href="static/PS/PS02_solutions.Rmd" download>`PS02_solutions.Rmd`</a> R Markdown "source code" 
* [`PS02_solutions.pdf`](static/PS/PS02_solutions.pdf){target="_blank"} PDF



***



# Problem set 01 {#PS01}

*Due Monday 9/13 5pm*

1. If you haven't already, complete the [intro survey](https://docs.google.com/forms/d/e/1FAIpQLSdA6uWnig5oR3ITCsVBFDxC6IC5Nq1tYzo-atdChe8CVTTqag/viewform){target="_blank"} distributed on Friday.
1. Complete a syllabus quiz.
    + Open the following [Google Doc](https://docs.google.com/document/d/1WknUhbtDeTGw8vvO8ytTURkF1_72ve4UXF5-SPk5NO8/){target="_blank"} and in the menu bar click "File -> Make a copy..."
    + Answer the questions based on information in the [syllabus](syllabus.html).
1. Submit a PDF of your syllabus quiz responses on gradescope:
    + Your gradescope account is based on your Smith, 5 College, or other official email.
    + Create your account on gradescope via the email sent to you this morning. If you did not receive such an email, add yourself to the course using Entry Code: 6PNXZ5.
    + Log into [gradescope](https://www.gradescope.com/courses/308310){target="_blank"} 
    + **Added after lecture**: You can edit gradescope submissions as many times as you like before the due date/time.
    + **Added after lecture**: You need to "tag" the 11 individual questions on whatever page(s) they are on (demo on Fri 9/10)
    



