---
title: "Your title here (REPLACE THIS)"
author: "Sec 01 AM, Group 76: Albert Y. Kim and P.K. Subban (REPLACE THIS)"
date: "Friday, November 22nd, 2019"
output:
  html_document:
    highlight: tango
    theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    df_print: kable
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Set R chunk default options:
# -do not output all R messages in html report
# -do not output all R warnings in html report
# -set default figure widths and heights for all code blocks, unless you 
#  override them
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2
)
```

```{r, echo=FALSE}
# Load all packages here. It's good practice to load all packages at the
# beginning of a Rmd file so that collaborators can see in one shot all packages
# they'll need installed:
library(tidyverse)
library(leaflet)
library(sf)
```

```{r, echo=FALSE}
# Load all the data here. Be sure to be in MP3 RStudio Project mode so that R
# knows where to find all data.
```



# Summary

Write no more than 100 words summarizing your work here.



# Introduction

Write no more than 250 words giving the data's context and source here.



# Static map section title 

Write no more than 250 words describing and interpreting your static map here. 

```{r, echo = FALSE}
# Put code to create your static map here:
ggplot() +
  geom_sf(data = bike_trails, col = "red")
```



# Interactive map section title 

Write no more than 250 words describing and interpreting your interactive map here. 

```{r, echo = FALSE}
# Put code to create your interactive map here:

```



# Additional Analysis

(Strictly optional) Add any additional material here.



# References and Citations

References should be generated using R Markdown footnotes, as you did in MP2.
